Tool-driven Revolutions in Archaeological Science

### 2018-12-27 analysis/paper/jcaa-paper.Rmd
white space
minor edits, what's wrong with some of the fig  captions?

In this paper we investigate recent developments which facilitate computational archaeology to return to principles of the scientific method and whether this might be the beginning of a paradigm-shift. Revolutions or paradigm-shifts in archaeology have been discussed to happen following theoretical discourses [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might not just be ideas-driven but also tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how Ecology and Archaeology have started to embrace methods which improve the reproducibility of their research. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code. A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology. We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology. Acknowledging this is not a simple process, we offer a R-based tool to facilitate the task of creating a compendium which enables other researchers to reproduce the published results.

Code and text for this paper are available at: https://github.com/benmarwick/caa2018-Tubingen

@@

@@

@@

```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency)), " words in ", prettyNum(nrow(archy_journal_articles)), " articles)")}

>>>>

In this paper we investigate recent developments which are facilitating computational archaeology's return to fundamental principles of the scientific method and whether this might be the beginning of a broader paradigm-shift. Revolutions or paradigm-shifts in archaeology have been proposed to occur following major theoretical statements [@Clark_1993; @Härke_2002], but there has been an argument in philosophy of science whether revolutions in science might  be not only ideas-driven, but also tool-driven [@Galison_1997; @Dyson_2000]. In this light we explore how the disciplines of Ecology and Archaeology have started to embrace tools that improve the reproducibility of their research. We focus on one set of tools for this change which is based on writing scripts in free and open source programming languages, exemplified by R, and the sharing of code. A review of bibliometric evidence showcases a strong development towards the use of R for Ecology and hints at the beginning of a similar development in archaeology. We thus evaluate how approaches to improving reproducibility and transparency in archaeology are mediated and transformed by digital approaches and propose these might be the sign of a tool-driven change in archaeology. Acknowledging this is not a simple process, we offer a R-based tool to facilitate the task of creating a compendium which enables other researchers to reproduce the published results.

@@

# Methods

@@

## Reproducibility and open source materials

To enable re-use of our materials and improve reproducibility and transparency according to the principles outlined in @Marwick2017, we include the entire R code used for all the analysis and visualizations contained in this paper at http://doi.org/10.17605/OSF.IO/RHVN5. Also in this version-controlled research compendium are the raw data for all the results reported here. All of the figures, tables and statistical test results presented here can be independently reproduced with the code and data in this repository. In our online materials our code is released under the MIT licence, our data as CC-0, and our figures as CC-BY, to enable maximum re-use [for more details, see @Marwick2017].

# Results

@@

```{r text-mining-article-titles-archaeology-plot, fig.cap=str_glue("Comparing the frequency of words used in titles \nof archaeology articles (", prettyNum(nrow(frequency), big.mark = ","), " words in ", prettyNum(nrow(archy_journal_articles), big.mark = ","), " articles)")}
